Reasoning About a Machine with Local Capabilities

### 2018-04-13 presentations/esop18-04-2018/main.tex
added awkward example as first example (probably want to use a different one)
presentaion

\begin{frame}
  \frametitle{Stack and return capabilities: Attack 1}
  \begin{tikzpicture}[scale=.5, every node={scale=.5}]
    % recurrent parts
    \stdstackstart

@@

    % stores stack pointer in heap
    \begin{onlyenv}<+->
      \begin{scope}
        \fill[gray!20] (14,6) rectangle (18,13);
        \draw (14,6) -- (14,13);
        \draw (18,6) -- (18,13);
      \end{scope}
      \draw (16,14) node {heap memory};
      \draw[fill=gray!50] (14,8) rectangle node[color=red] {\scriptsize copy of old sp} (18,8.5);
    \end{onlyenv}
    \draw<.> (14,8.25) edge[->,red,thick,bend left] (sp1);
    \begin{onlyenv}<+->
      \draw [decorate,decoration={brace,amplitude=10pt,mirror,raise=4pt},yshift=0pt]
      (11.5,6.5) -- (11.5,16) node[draw=black] (spold) [black,midway,xshift=0.8cm] {};
      \draw (14,8.25) edge[->,red,thick,bend left] (spold);
    \end{onlyenv}

    % attacker returns
    \begin{onlyenv}<+>
      \draw [decorate,decoration={brace,amplitude=10pt,mirror,raise=4pt},yshift=0pt]
      (6.5,2.5) -- (6.5,16) node[draw=black] (sp1) [black,midway,xshift=0.8cm] {};
      \draw node[right=0cm of sp1] { stack pointer };
      \draw[fill=white] (0,2.5) rectangle (6,6.5) node[pos=.5] {\footnotesize current stack frame};
    \end{onlyenv}

    % attacker gets called again + uses old stack pointer.
    \begin{onlyenv}<+>
      \draw[fill=gray!50] (0,2.5) rectangle (6,10.5) node[pos=.5] {\footnotesize caller stack frame};
      \actadv{(0,10.5)}{(6,13.5)}{callee stack frame}
      \draw (14,8.25) edge[->,color=red,very thick,out=235,in=0] node[sloped, below] {read/write} (6,8);
      \draw [decorate,decoration={brace,amplitude=10pt,mirror,raise=4pt},yshift=0pt]
        (6.5,10.5) -- (6.5,16) node[draw=black] (sp1) [black,midway,xshift=0.8cm] {};
      \draw node[right=0cm of sp1] { stack pointer };
      % \draw [decorate,decoration={brace,amplitude=10pt,mirror,raise=4pt},yshift=0pt]
      % (6,2.5) -- (6,16) node (callerframe1) [draw=black,midway,xshift=0.8cm] {};
      % \draw node[fill=gray!50,right=0cm of callerframe1] (rp1l) {return pointer data};
    \end{onlyenv}
  \end{tikzpicture}
\end{frame}

\begin{frame}
  \frametitle{Stack and return capabilities: Attack 2}
  \begin{tikzpicture}[scale=.5, every node={scale=.5}]
    % recurrent parts
    \stdstackstart

@@

    % attacker uses old stack pointer
    \draw<+->[fill=gray!50] (0,7) rectangle (6,9) node[pos=.5] {\footnotesize callee (2) stack frame};
    \begin{onlyenv}<.-.(1)>
      \actadv{(0,9)}{(6,12)}{callee (3) stack frame}
      \draw [decorate,decoration={brace,amplitude=10pt,mirror,raise=4pt},yshift=0pt]
      (6.5,9) -- (6.5,16) node[draw=black] (sp1) [black,midway,xshift=0.8cm] {};
      \draw node[right=0cm of sp1] { stack pointer };
    \end{onlyenv}
    \draw<+> (spold) edge[->,color=red,very thick,out=235,in=0] node[sloped, below] {read/write} (6,8);
  \end{tikzpicture}
\end{frame}

\begin{frame}
  \begin{tikzpicture}[scale=.5, every node={scale=.5}]
    \frametitle{Stack and return capabilities: Attack 3}
    \begin{scope}
    \clip (-.1,-.1) rectangle (5.1,10.1);
    \fill[fill=gray!50] (0,0) rectangle (5,10);
    \draw (0,0) -- (0,10);
    \draw (5,0) -- (5,10);
    \end{scope}
    \draw (2.5,11) node {code memory};

    \begin{scope}
    \clip (12.9,3.9) rectangle (18.1,11.1);
    \fill[fill=gray!50] (13,4) rectangle (18,11);
    \draw (13,4) -- (13,11);
    \draw (18,4) -- (18,11);
    \end{scope}
    \draw (15.5,12) node {more code memory};

    \begin{scope}
    \clip (12.9,-.1) rectangle (18.1,2.1);
    \fill[fill=gray!50] (13,0) rectangle (18,2);
    \draw (13,0) -- (13,2);
    \draw (18,0) -- (18,2);
    \end{scope}
    \draw (14.5,3) node {heap memory};

\note{Caller calls callee}
    \draw<+->[fill=white] (0,1) rectangle (5,4) node[pos=.5] {\footnotesize caller code};

    \begin{onlyenv}<+->
      \capbrace[ret1]{(5.1,1)}{(5.1,4)}
      \draw node[fill=gray!50, right=0cm of ret1] { \phantom{ret-ptr} };
      \draw node[right=0cm of ret1] { ret-ptr };
    \end{onlyenv}

\note{Callee has access to some place to store things.}
    \begin{onlyenv}<+->
      \draw[fill=gray!30] (0,1) rectangle (5,4) node[pos=.5] {\footnotesize caller code};
      \draw[fill=white] (13,5) rectangle (18,6) node[pos=.5] {\footnotesize callee code};
      \draw[fill=white] (13,1) rectangle (18,1.5) node[pos=.5] {};
    \end{onlyenv}

    \begin{onlyenv}<.-.(2)>
      \draw [decorate,decoration={brace,amplitude=2pt,mirror,raise=4pt},yshift=0pt]
      (18.1,1) -- (18.1,1.5) node[draw=black] (copy) [black,midway,xshift=0.5cm] {};
      \draw (18,5.5)  edge[->,bend left] (copy);
    \end{onlyenv}

\note{Callee stores the return pointer they got from caller.}
    \begin{onlyenv}<+->
      \draw[fill=white] (13,1) rectangle (18,1.5) node[pos=.5] {\scriptsize {\color{red} copy of ret-ptr}};
      \draw (13,1.25) edge[->,red,thick,bend left] (ret1);
    \end{onlyenv}

\note{Callee calls callee (1)}
    \begin{onlyenv}<+->
      \draw [decorate,decoration={brace,amplitude=3pt,raise=4pt},yshift=0pt]
      (12.9,5) -- (12.9,6) node[draw=black] (ret2) [black,midway,xshift=-0.8cm] {};
      \draw node[fill=gray!50,left=0cm of ret2] { \phantom{ret-ptr (1)} };
      \draw node[left=0cm of ret2] (retptr1) { ret-ptr (1) };
    \end{onlyenv}

    \begin{onlyenv}<+->
      \draw[fill=white] (0,5) rectangle (5,8) node[pos=.5] {\footnotesize callee code (1)};
      \draw[fill=gray!30] (13,5) rectangle (18,6) node[pos=.5] {\footnotesize callee code};
      \draw[fill=gray!30] (13,1) rectangle (18,1.5) node[pos=.5] {\scriptsize {\color{red} copy of ret-ptr}};
    \end{onlyenv}

\note{Callee (1) calls callee (2)}
    \begin{onlyenv}<+->
      \capbrace[ret3]{(5.1,5)}{(5.1,8)}
      \draw node[fill=gray!50,right=0cm of ret3] { \phantom{ret-ptr (2)} };
      \draw node[right=0cm of ret3] (retptr2) { ret-ptr (2) };
    \end{onlyenv}

    \begin{onlyenv}<+->
      \draw[fill=gray!30] (0,5) rectangle (5,8) node[pos=.5] {\footnotesize callee code (1)};
      \draw[fill=white] (13,7) rectangle (18,10) node[pos=.5] {\footnotesize callee code (2)};
    \end{onlyenv}
\note{Callee (2) colludes with callee}
    \begin{onlyenv}<+->
      \draw (21,7.5) node[devil,opaque=0.5,mirrored,minimum size=1cm] {};
    \end{onlyenv}
\note{Callee (2) has access to ret-ptr (the return pointer caller passed to callee)}
    \begin{onlyenv}<+->
      \draw [decorate,decoration={brace,amplitude=2pt,mirror,raise=4pt},yshift=0pt]
      (18.1,1) -- (18.1,1.5) node[draw=black] (copy) [black,midway,xshift=0.5cm] {};
      \draw (18,8.25)  edge[->,red,thick,bend left] (copy);
      \draw[fill=white] (13,1) rectangle (18,1.5) node[pos=.5] {\scriptsize {\color{red} copy of ret-ptr}};
    \end{onlyenv}

\note{Callee (2) uses ret-ptr to return}
    \begin{onlyenv}<+->
      \draw[fill=gray!30] (13,7) rectangle (18,10) node[pos=.5] {\footnotesize callee code (2)};
      \draw[fill=white] (0,1) rectangle (5,4) node[pos=.5] {\footnotesize caller code};
      \draw[fill=gray!30] (13,1) rectangle (18,1.5) node[pos=.5] {\scriptsize {\color{red} copy of ret-ptr}};
    \end{onlyenv}

\note{This breaks well-bracketedness as ret-ptr (2) and ret-ptr (1) was not returned to}
    \begin{onlyenv}<+->
      \draw node [ellipse, fit=(retptr1) (retptr2), draw, red] { };
      \draw node[red,cloud,cloud puffs=10.8,cloud puff arc=110, aspect=3, draw, fill=white, align=center] () at (12,10) { These were skipped!};  
    \end{onlyenv}
  \end{tikzpicture}
\end{frame}
%%% Local Variables:
%%% TeX-master: "presentation-illustrations"
%%% End:

>>>>

\documentclass{beamer}
\addtobeamertemplate{navigation symbols}{}{%
    \usebeamerfont{footline}%
    \usebeamercolor[fg]{footline}%
    \hspace{1em}%
    \insertframenumber/\inserttotalframenumber
  }

\usepackage{listings}
\renewcommand{\ttdefault}{pcr}
\lstset{
  basicstyle=\ttfamily,
  keywordstyle=\ttfamily\bfseries,
  morekeywords={let, ref, assert, in},
  columns=fullflexible,
  keepspaces=true,
  escapeinside={(*}{*)}
}

\input{../illustrations/preamble}
\input{loc_preamble}
\title{Reasoning About a Machine with Local Capabilities}
\subtitle{Provably Safe Stack and Return Pointer Management}
\author{Lau Skorstengaard\inst{1} \and Dominique Devriese\inst{2} \and Lars Birkedal\inst{1}}
\institute{\inst{1} Aarhus University \and %
  \inst{2} imec-DistriNet, KU Leuven}

\date{ESOP, April 17, 2018}

\tikzset{
    gray on/.style={alt=#1{gray}{}},
    alt/.code args={<#1>#2#3}{%
      \alt<#1>{\pgfkeysalso{#2}}{\pgfkeysalso{#3}} % \pgfkeysalso doesn't change the path
    },
}

\begin{document}
\maketitle

\begin{frame}[fragile]
  \frametitle{How Do We Reason About Programs Informally}
  % Motivation
  \centering

  \begin{tabular}{c}
    \begin{lstlisting}
let x = ref 0 in
  (*$\lambda$*)f.(x := 0;
      f();
      x := 1;
      f();
      assert(x == 1))
    \end{lstlisting}
  \end{tabular}

  \note{Informally, what does this program do?}
  \note{It allocates a reference to 0 and saves a closure to x. The closure takes a callback, assigns 0 to x, calls the callback, assigns 1 to x, calls the callback another time and finally asserts x to be 1.}
  \note{As a programmer that is how I would think. But what happens if I compile this closure to assembly and let some untrusted piece of code interact with it?}
  \note{My reasoning depends on the assumption that when I call f, then if f returns it returns to a certain point. The low-level machine needs to enforce this if I want to be able to do have this closure interact with arbitrary machine code.} 
\end{frame}

\begin{frame}
  \frametitle{Traditional Stack Pointers}
  \note{Let's first consider what happens with the stack during a call on a traditional low-level machine and what can go wrong.}
  \begin{tikzpicture}[scale=.5, every node={scale=.5}]
    % recurrent parts
    \begin{scope}
      \clip (-.1,-.1) rectangle (6.1,15.1);
      \fill[gray!20,draw=none,opacity=.8] (0,0) rectangle (6,15);
      \draw[draw=gray!80] (0,0) -- (0,15);
      \draw[draw=gray!80] (6,0) -- (6,15);
      \draw[fill=white] (0,-.5) rectangle (6,2.5) node[pos=.5] {\footnotesize lower stack frames...};
    \end{scope}
    \draw (-1.5,0) node {};
    \draw (-1.5,15) node {};
    % \draw[->] (-1,0) -- node[midway,sloped,above] {stack grows upward} (-1,15);

    % traditional:
    \draw<+>[fill=white] (0,2.5) rectangle (6,5.5) node[pos=.5] {\footnotesize caller stack frame};
    \draw<.> (8,7) node[draw=black] (sp0) {};
    \draw<.> node[right=0cm of sp0] (sp0l) { stack pointer };
    \draw<.> (sp0) edge[->,out=235,in=0] (6,5.25);

    % traditional call:
    \draw<+>[fill=white] (0,2.5) rectangle (6,5.5) node[pos=.5] {\footnotesize caller stack frame};
    \draw<.>[fill=white] (0,5.5) rectangle (6,8.5) node[pos=.5] {\footnotesize callee stack frame};
    \draw<.> (8,7) node[draw=black] (sp0) {};
    \draw<.> node[right=0cm of sp0] (sp0l) { stack pointer };
    \draw<.> (sp0) edge[->,out=235,in=0] (6,8.25);

    % traditional return:
    \draw<+>[fill=white] (0,2.5) rectangle (6,5.5) node[pos=.5] {\footnotesize caller stack frame};
    \draw<.> (8,7) node[draw=black] (sp0) {};
    \draw<.> node[right=0cm of sp0] (sp0l) { stack pointer };
    \draw<.> (sp0) edge[->,out=235,in=0] (6,5.25);

    % traditional call, attack:
    \draw<+->[fill=white] (0,2.5) rectangle (6,5.5) node[pos=.5] {\footnotesize caller stack frame};
    \draw<.->[fill=white] (0,5.5) rectangle (6,8.5) node[pos=.5] {\footnotesize callee stack frame};
    \draw<.-> (8,7) node[draw=black] (sp0) {};
    \draw<.-> node[right=0cm of sp0] (sp0l) { stack pointer };

    % read other stack frames
    \draw<.> (sp0) edge[->,out=235,in=0] (6,8.25);
    \actadv[.]{(0,5.5)}{(6,8.5)}{callee stack frame}
    \draw<.> (sp0) edge[->,color=red,very thick,out=235,in=0] node[sloped, below] {read/write} (6,4);

    % break well-bracketedness
    \draw<+>[fill=red,opacity=.7] (-.5,2.5) rectangle (6.5,8.5);
    \draw<.-> (8,7) node[draw=black] (sp0) {};
    \draw<.>[red] (sp0) edge[->,out=235,in=0] (6,1.75);
    \draw<.> node[color=red,align=left,below=of sp0l] {return but\\
      skip caller frame};
  \end{tikzpicture}
  \note{We need something to enforce security properties on a low-level machine.}
\end{frame}

\begin{frame}
  \frametitle{Capability Machine}
  % short description of capability machines

  \begin{columns}
\begin{column}{0.5\textwidth}
  \begin{itemize}
  \item Low-level machine
    \note{Has instructions as you would expect, load, store, jmp, etc.}
  \item Capabilities replace pointers
    \begin{itemize}
    \item Pointer
    \item Range of authority
    \item Kind of authority
      \begin{itemize}
      \item read/write/execute
      \item enter
      \end{itemize}
      \note{Has instructions for manipulating capabilities.}
    \end{itemize}
    \item Authority checked dynamically
  \end{itemize}
\end{column}
\begin{column}{0.5\textwidth}
    \begin{center}
      % Insert capability figure
     \end{center}
\end{column}
\end{columns}
  \note{This particular setup gives a very fine-grained memory control.}
\end{frame}

\begin{frame}
  \frametitle{Stack and Return Capabilities: Attack 1}
  \note{In the following, part of memory is used for the stack and we use capabilities to govern them. Callee not able to read from the caller stack frame, so that is already an improvement.}
  \begin{tikzpicture}[scale=.5, every node={scale=.5}]
    % recurrent parts
    \stdstackstart

@@

    \note{Evil callee stores stack pointer on the heap.}
    % stores stack pointer in heap
    \begin{onlyenv}<+->
      \begin{scope}
        \fill[gray!20] (14,6) rectangle (18,13);
        \draw (14,6) -- (14,13);
        \draw (18,6) -- (18,13);
      \end{scope}
      \draw (16,14) node {heap memory};
      \draw[fill=gray!50] (14,8) rectangle node[color=red] {\scriptsize copy of old sp} (18,8.5);
    \end{onlyenv}
    \draw<.> (14,8.25) edge[->,red,thick,bend left] (sp1);
    \begin{onlyenv}<+->
      \draw [decorate,decoration={brace,amplitude=10pt,mirror,raise=4pt},yshift=0pt]
      (11.5,6.5) -- (11.5,16) node[draw=black] (spold) [black,midway,xshift=0.8cm] {};
      \draw (14,8.25) edge[->,red,thick,bend left] (spold);
    \end{onlyenv}

    \note{Evil callee returns to the caller.}
    % attacker returns
    \begin{onlyenv}<+>
      \draw [decorate,decoration={brace,amplitude=10pt,mirror,raise=4pt},yshift=0pt]
      (6.5,2.5) -- (6.5,16) node[draw=black] (sp1) [black,midway,xshift=0.8cm] {};
      \draw node[right=0cm of sp1] { stack pointer };
      \draw[fill=white] (0,2.5) rectangle (6,6.5) node[pos=.5] {\footnotesize current stack frame};
    \end{onlyenv}

    \note{caller uses a bit more of the stack and calls the evil callee again.}
    \note{the evil callee can now load the old stack capability from memory and access part of the callers private stack!}
    % attacker gets called again + uses old stack pointer.
    \begin{onlyenv}<+>
      \draw[fill=gray!50] (0,2.5) rectangle (6,10.5) node[pos=.5] {\footnotesize caller stack frame};
      \actadv{(0,10.5)}{(6,13.5)}{callee stack frame}
      \draw (14,8.25) edge[->,color=red,very thick,out=235,in=0] node[sloped, below] {read/write} (6,8);
      \draw [decorate,decoration={brace,amplitude=10pt,mirror,raise=4pt},yshift=0pt]
        (6.5,10.5) -- (6.5,16) node[draw=black] (sp1) [black,midway,xshift=0.8cm] {};
      \draw node[right=0cm of sp1] { stack pointer };
      % \draw [decorate,decoration={brace,amplitude=10pt,mirror,raise=4pt},yshift=0pt]
      % (6,2.5) -- (6,16) node (callerframe1) [draw=black,midway,xshift=0.8cm] {};
      % \draw node[fill=gray!50,right=0cm of callerframe1] (rp1l) {return pointer data};
    \end{onlyenv}
  \end{tikzpicture}
\end{frame}

\begin{frame}
  \frametitle{Local Capabilities}
  % Short description of local capabilities
  \begin{itemize}
  \item Capabilities tagged with locality (local or global)   \note{We call non-local capabilities global}
  \item New \emph{write-local} permission.
  \item Local capabilities can only be stored by capabilities with \emph{write-local} permission
  \end{itemize}
  Calling convention highlights
  \begin{itemize}
  \item Stack capability is local with permission read, write-local, and execute.
  \item Clear stack before passing stack capability to untrusted code.
  \end{itemize}
\end{frame}

\begin{frame}
  \frametitle{Local Stack Capabilities Prevent Attack 1}
  \begin{tikzpicture}[scale=.5, every node={scale=.5}]
    % recurrent parts
    \stdstackstart

    % attacker...
    \draw[fill=gray!50] (0,2.5) rectangle (6,6.5) node[pos=.5] {\footnotesize caller stack frame};
    \actadv{(0,6.5)}{(6,9.5)}{callee stack frame}
    \draw [decorate,decoration={brace,amplitude=10pt,mirror,raise=4pt},yshift=0pt]
      (6.5,6.5) -- (6.5,16) node[draw=black] (sp1) [black,midway,xshift=0.8cm] {};
    \draw node[right=0cm of sp1] { stack pointer };

    % stores stack pointer in heap
    \begin{scope}
      \fill[gray!20] (14,6) rectangle (18,13);
      \draw (14,6) -- (14,13);
      \draw (18,6) -- (18,13);
    \end{scope}
    \draw (16,14) node {heap memory};
    \draw[fill=gray!50] (14,8) rectangle node[color=red] {\scriptsize copy of old sp} (18,8.5);
    \draw (14,8.25) edge[->,red,thick,bend left] (sp1);
    \draw[teal,very thick] (13.5,7.5) -- (18.5,9);
    \draw[teal,very thick] (18.5,7.5) -- (13.5,9);
    \draw node[teal,cloud,cloud puffs=10.8,cloud puff arc=110, aspect=3, draw, fill=white] () at (14,4) {Stack pointer is local!};  
\end{tikzpicture}
\end{frame}

\begin{frame}
  \frametitle{Stack and Return Capabilities: Attack 2}
  \begin{tikzpicture}[scale=.5, every node={scale=.5}]
    % recurrent parts
    \stdstackstart

@@

    % attacker uses old stack pointer
    \draw<+->[fill=gray!50] (0,7) rectangle (6,9) node[pos=.5] {\footnotesize callee (2) stack frame};
    \begin{onlyenv}<.-.(1)>
      \actadv{(0,9)}{(6,12)}{callee (3) stack frame}
      \draw [decorate,decoration={brace,amplitude=10pt,mirror,raise=4pt},yshift=0pt]
      (6.5,9) -- (6.5,16) node[draw=black] (sp1) [black,midway,xshift=0.8cm] {};
      \draw node[right=0cm of sp1] { stack pointer };
    \end{onlyenv}
    \draw<+> (spold) edge[->,color=red,very thick,out=235,in=0] node[sloped, below] {read/write} (6,8);
  \end{tikzpicture}
  \note{The stack was the only place to store the local capability, so the adversary hid it there.}
\end{frame}

\begin{frame}
  \frametitle{Calling Convention (Continued)}
  \begin{itemize}
  \item $\dots$
  \item Clear stack and non-argument registers before invoking untrusted code.
  \end{itemize}
\end{frame}

\begin{frame}
  \frametitle{Stack Clearing Prevents Attack 2}
  \begin{tikzpicture}[scale=.5, every node={scale=.5}]
    % recurrent parts
    \stdstackstart
    % attacker stores copy of stack pointer high in the stack
    \draw[fill=gray!50] (0,2.5) rectangle (6,4) node[pos=.5] {\footnotesize caller stack frame};
    \begin{onlyenv}<+>
      \actadv{(0,4)}{(6,7)}{callee stack frame}
    \end{onlyenv}
    \draw<.-.(2)>[fill=white] (0,13) rectangle (6,13.5) node[pos=.5,color=red] {\footnotesize copy of sp};
    \draw<.-.(2)> [decorate,decoration={brace,amplitude=10pt,mirror,raise=4pt},yshift=0pt]
    (13.5,4) -- (13.5,16) node[draw=black] (spold) [black,midway,xshift=0.8cm] {};
    \draw<.-.(2)> (6,13.25) edge[red,thick,->,in=45] (spold);

    % attacker calls trusted code again
    \begin{onlyenv}<+->
      \inactadv{(0,4)}{(6,7)}{callee stack frame}
    \end{onlyenv}
    \begin{onlyenv}<.-.(1)>
      \draw[fill=white] (0,7) rectangle (6,9) node[pos=.5] {\footnotesize callee (2) stack frame};
      \draw [decorate,decoration={brace,amplitude=10pt,mirror,raise=4pt},yshift=0pt]
        (6.5,7) -- (6.5,16) node[draw=black] (sp1) [black,midway,xshift=0.8cm] {};
      \draw node[right=0cm of sp1] { stack pointer };
    \end{onlyenv}

    % trusted code clears the stack
    \begin{onlyenv}<+->
      \foreach \x in {9,9.5,...,14}
      {
        \draw[fill=white] (0,\x) rectangle (6,\x+.5) node[pos=.5,color=teal] {\footnotesize 0};
      };
    \end{onlyenv}
    \draw<.>[very thick,color=teal] (10,13.5) -- (13,14.5)
    (13,13.5) -- (10,14.5);

    % attacker uses old stack pointer
    \draw<+->[fill=gray!50] (0,7) rectangle (6,9) node[pos=.5] {\footnotesize callee (2) stack frame};
    \begin{onlyenv}<.>
      \actadv{(0,9)}{(6,12)}{callee (3) stack frame}
      \draw [decorate,decoration={brace,amplitude=10pt,mirror,raise=4pt},yshift=0pt]
      (6.5,9) -- (6.5,16) node[draw=black] (sp1) [black,midway,xshift=0.8cm] {};
      \draw node[right=0cm of sp1] { stack pointer };
    \end{onlyenv}

    % attacker returns
    \draw<+>[fill=white] (0,9) rectangle (6,12) node[pos=.5] {?};
    \begin{onlyenv}<.>
      \draw[fill=white] (0,7) rectangle (6,9) node[pos=.5] {\footnotesize callee (2) stack frame};
      \draw [decorate,decoration={brace,amplitude=10pt,mirror,raise=4pt},yshift=0pt]
        (6.5,7) -- (6.5,16) node[draw=black] (sp1) [black,midway,xshift=0.8cm] {};
      \draw node[right=0cm of sp1] { stack pointer };
    \end{onlyenv}

    % trusted code clears stack again
    \begin{onlyenv}<+->
      \foreach \x in {7,7.5,...,14}
      {
        \draw[fill=white] (0,\x) rectangle (6,\x+.5) node[pos=.5,color=teal] {\footnotesize 0};
      };
    \end{onlyenv}
    % then, the trusted code can return
    \begin{onlyenv}<+>
      \actadv{(0,4)}{(6,7)}{callee stack frame}
    \end{onlyenv}

  \end{tikzpicture}
\end{frame}

\begin{frame}
  \frametitle{(Full) Calling Convention}
  % and stack management
  \note{We need to do a few more things, but this is the rest (just to illustrate that it is fairly simple).}
  \note{Every point is motivated by some attack.}
  \begin{itemize}
  \item Initially:
    \begin{itemize}
    \item Stack capabilitiy local capability with read, write-local, and execute authority.
    \item No global write-local capabilities on the machine.
    \end{itemize}
  \item Prior to returning to untrusted code:
    \begin{itemize}
    \item Clear the stack.
    \item Clear non-return registers.
    \end{itemize}
  \item Prior to calls to untrusted code:
    \begin{itemize}
    \item Push activation record to the stack and create enter-capability.
    \item Restrict the stack pointer to the unused part and clear that part.
    \item Clear non-argument registers.
    \end{itemize}
  \item Only invoke global call-backs.
  \item When invoked by untrusted code
    \begin{itemize}
    \item Make sure the stack pointer has read, write-local and execute authority.
    \end{itemize}
  \end{itemize}
\end{frame}

\begin{frame}
  \frametitle{Formalizing the Guarantees of a Capability Machine}
  \note{How can we be sure calling convention works. Specifically, if a program interacts with intrusted code using the CC, can we formally show the correctness of the program if it relies on well-bracketedness or local-state encapsulation.}

  \note{Need formal statement of the guarantees provided by the capability machine including the specific guarantees for local capabilities.}
  \begin{itemize}
  \item How do we know the calling convention works?
  \item Unary step-indexed Kripke logical relation over recursive worlds
    \begin{itemize}
    \item Statement of guarantees probided by the capability machine
    \end{itemize}
  \end{itemize}
  \note{We state this formal statement in terms of a unary step-indexed Kripke logical relation over recursive worlds}

  \note{Calling convention main application, but it is very general - can be used to reason about other programs.}

  \note{In the following: give some intuition for different parts of LR}

  \note{correctness here could be assert not violated}

  \note{mention better than previous. Define the guarantees}
\end{frame}

\begin{frame}
  \frametitle{Worlds, Safe Values, and Step-Indexing}
\begin{columns}
  \begin{column}{0.5\textwidth}
    \begin{itemize}
    \item<1-> Capabilities represent bound on executing code
    \item<2-> World, $W$      
      \begin{itemize}
      \item Collection of invariants
      \end{itemize}
    \item<4-> Predicate for safe values w.r.t world, $\stdvr(W)$
      \begin{itemize}[<8->]
      \item Recursively defined
      \end{itemize}
    \end{itemize}
\end{column}
\begin{column}{0.5\textwidth}
  \begin{tikzpicture}[scale=.5, every node={scale=.5}]
    \scope
      \clip (-.1,-.1) rectangle (6.1,15.1);
      \fill[fill=white] (0,0) rectangle (6,15);
      \draw (0,0) -- (0,15);
      \draw (6,0) -- (6,15);
    \endscope
    % \draw[->] (-2,0) -- node[midway,sloped,above] {stack grows upward} (-2,15);
    \draw (-1.5,0) node {};
    \draw (-1.5,15) node {};
    \foreach \x in {0.5,1,...,14}
    {
      \draw[fill=white] (0,\x) rectangle (6,\x+.5) node[pos=.5,color=teal] {};
    };

    % One capability
    \draw<1-6> [decorate,decoration={brace,amplitude=3pt,mirror,raise=-2pt},yshift=0pt]
    (6.25,2) -- (6.25,5) node[draw=black] (sp1) [black,midway,xshift=6pt] {}; % ,shape=coordinate,xshift=1pt
    \draw<1-4>  (6,7.25) edge[black,->,in=45] (sp1);

    % Invariant colored
    \fill<2->[red,opacity=.7] (0,1) rectangle (6,6);
    \fill<2->[blue,opacity=.7] (0,7) rectangle (6,10);

    \fill<3-6>[black, fill opacity=0, text opacity=0.7] (0,4) rectangle (6,4.5) node[pos=.5] {\footnotesize 42};

    \draw<5> (6,7.25) edge[black,->,in=45] node[right] {\footnotesize \textsc{r}} (sp1)   ;
    \draw<6> (6,7.25) edge[black,->,in=45] node[right] {\footnotesize \textsc{rw}} (sp1)   ;
    \draw<6>[very thick,color=red] (6.5,6.5) -- (8.5,4.5)
    (6.5,4.5) -- (8.5,6.5);

    \draw<7-> [decorate,decoration={brace,amplitude=3pt,mirror,raise=-2pt},yshift=0pt]
    (6.25,2) -- (6.25,5) node[draw=black] (sp1) [black,midway,xshift=6pt] {}; % ,shape=coordinate,xshift=1pt
    \path<7->  (6,4.25) edge[black,out=30,in=135] node[above right]  {\footnotesize \textsc{r}} (8,5) 
              (8,5) edge[black,->,out=-45,in=45](sp1);

  \end{tikzpicture}

\end{column}
\end{columns}
  % Do this as a series of figures?

  % Figure with memory and one capability to make the bound explicit
  \note{compared to normal assembly, capabilities represent bounds on executing block of code.}

  \note{we have no observable I/O, so the authority bounds we consider are related to memory. However, more fine-grained/detailed than read/write-authority.}

  % Color the memory to illustrate different regions of the world
  \note{a piece of code can be bound by arbitrary memory invariants which we define in a \emph{world}.}

  % Write down a static invariant (some address contains 42, maybe write a legend on the left side of slide.).
  \note{essentially, a world is a collection of invariants and safety of words defined with respect to a world.}

  \note{Define a set of words that are safe w.r.t. world W V(W) in P(Word)}

  \note{Whether a capability is safe depends on the authority it carries}

  % Add a read capability to the address that contains 42, safe
  % Add read/write capability to the address that contains 42, not safe - can be used to violate the invariant.
  \note{Example, w.r.t. a world with the invariant that an address contains constant. Safe for read capability, not write as write capability can break this invariant.}

  % Add a read capability that refers to itself
  \note{safety for a read capability is the case if the read capability only gives access to safe capabilities.}

  \note{What if that part of memory contains a read capability for the same part of memory? Cyclic definition. Solved by step-indexing.}

  \note{related to similar issue with languages with recursive types or higher-order ML-list references.}

  \note{Solved by step-indexing - safety up to a certain number of interaction steps.}
\end{frame}

\begin{frame}
  \frametitle{Future Worlds and Invariants, and Recursive Worlds}
  \begin{tikzpicture}[scale=.5, every node={scale=.5}]
    \scope
      \clip (-.1,-.1) rectangle (6.1,8.1);
      \fill[fill=white] (0,0) rectangle (6,8);
      \draw (0,0) -- (0,8);
      \draw (6,0) -- (6,8);
    \endscope
    % \draw[->] (-2,0) -- node[midway,sloped,above] {stack grows upward} (-2,15);
    \draw (-1.5,0) node {};
    \draw (-1.5,8) node {};
    \foreach \x in {0.5,1,...,7}
    {
      \draw[fill=white] (0,\x) rectangle (6,\x+.5) node[pos=.5,color=teal] {};
    };

    \fill[red,opacity=.7] (0,7.5) rectangle (6,6);
    \fill[blue,opacity=.7] (0,6) rectangle (6,4);

    \begin{onlyenv}<2->
    \scope
      \clip (11.9,-.1) rectangle (18.1,8.1);
      \fill[fill=white] (12,0) rectangle (18,8);
      \draw (12,0) -- (12,8);
      \draw (18,0) -- (18,8);
    \endscope
    % \draw[->] (-2,0) -- node[midway,sloped,above] {stack grows upward} (-2,15);
    \draw (-1.5,0) node {};
    \draw (-1.5,8) node {};
    \foreach \x in {0.5,1,...,7}
    {
      \draw[fill=white] (12,\x) rectangle (18,\x+.5) node[pos=.5,color=teal] {};
    };

    \fill[red,opacity=.7] (12,7.5) rectangle (18,6);
    \fill[blue,opacity=.7] (12,6) rectangle (18,4);

    \draw [decorate,decoration={brace,amplitude=3pt,raise=2pt},yshift=0pt]
    (11.75,1) -- (11.75,4) node[draw=black] (sp1) [black,midway,xshift=-12pt] {}; % ,shape=coordinate,xshift=1pt
    \draw  (10,0.5) edge[black,->,in=225] (sp1);

    \draw (7,7.5) edge [black,thick, ->] node[above] {\footnotesize Memory allocated} (11,7.5);
  \end{onlyenv}
  \begin{onlyenv}<3->
    \fill[green,opacity=.7] (12,4) rectangle (18,1);

    \draw (9,4) node {\huge $\sqsubseteq$};
  \end{onlyenv}
    %Invariant transition systems
    \begin{onlyenv}<4->
    \node[circle,draw=black, fill=white] (a) at (3,6.75) {};
    \draw[>=stealth] (2,6.75) edge[black,->] (a);

    \node[circle,draw=black, fill=white] (b) at (2,5) {};
    \draw[>=stealth] (1,5) edge[black,->] (b);
    \node[circle,draw=black, fill=white] (c) at (4,5) {};
    \draw[>=stealth] (b) edge[black,->, bend left] (c);
    \draw[>=stealth] (c) edge[black,->, bend left] (b);

    \node[circle,draw=black, fill=white] (a2) at (15,6.75) {};
    \draw[>=stealth] (14,6.75) edge[black,->] (a2);

    \node[circle,draw=black, fill=white] (b2) at (14,5) {};
    \draw[>=stealth] (13,5) edge[black,->] (b2);
    \node[circle,draw=black, fill=white] (c2) at (16,5) {};
    \draw[>=stealth] (b2) edge[black,->, bend left] (c2);
    \draw[>=stealth] (c2) edge[black,->, bend left] (b2);

    \node[circle,draw=black, fill=white] (d) at (14,2.5) {};
    \draw[>=stealth] (13,2.5) edge[black,->] (d);
    \node[circle,draw=black, fill=white] (e) at (16,3.25) {};
    \draw[>=stealth] (d) edge[black,->, bend left] (e);
    \draw[>=stealth] (e) edge[black,->, bend left] (d);
    \node[circle,draw=black, fill=white] (f) at (16,1.75) {};
    \draw[>=stealth] (e) edge[black,->, bend left] (f);
  \end{onlyenv}
\end{tikzpicture}
